# Social Desirability Bias {#sec-desirability}

```{r, include = FALSE}
library(lavaan)
```

Much of the research carried out with human beings that measures behaviors, affections, personality, etc. use self-report scales (Lange & Dewitte, 2019; Peterson & Kerin, 1981). When responding to a questionnaire, some factors influence the response given to items that may or may not be associated with the latent trait being measured. Ideally, when we measure a construct, we want to measure it without many errors or spurious variations; however, it is possible that there is bias/response style that introduces spurious variations into our analyses. Some examples of these responses are: social desirability, acquiescence, and extreme responses.

## Faking: The Good, The Bad, and the Ugly

*Faking* depends on the context of the application and the questionnaire applied. The person who uses *faking* aims to provide a representation of themselves that helps achieve a personal objective (Ziegler et al., 2011). Therefore, *faking* occurs when this set of responses is activated by situational demands and personal characteristics to produce systematic differences in test scores that are not due to the construct of interest. *Faking* is a behavior that is influenced by different factors and is, in essence, a matter of measurement (Ziegler et al., 2011).

*Faking* can be conceptualized as *faking good* and *faking bad*. *Faking good* is a conscious effort to manipulate responses to an instrument to make a positive impression (Zickar & Robie, 1999). *Faking bad* includes both the fabrication of clinical and/or diagnostic symptoms and the exaggeration of symptoms to obtain a specific secondary gain (Ziegler et al., 2011). One question that remains is: what makes people pretend?

Variables in *faking* models can be classified based on the type of belief a given variable is likely to impact. The expectancy theory of Ziegler et al. (2011) states that the choice to do *faking* or not is caused by: a) Belief that one is capable of doing *faking*; b) Belief that doing *faking* is important; c) Belief that the opportunity is valued. The belief that someone is capable of *faking* comes from different variables, such as personality traits, cognitive ability, knowledge and experience, as well as situational factors such as the degree of transparency of the item and the use of verification warnings that make an individual more or less capable of faking (Griffith et al., 2006; McFarland & Ryan, 2000; Raymark & Tafero, 2009; Riggio et al., 1988; Snell et al., 1999).

## Faking Good: Social Desirability

A comprehensive survey of existing literature indicates that Social Desirability (SD) scales stand out as the most frequently utilized and explored measures of faking behavior (Ziegler et al., 2011). When individuals are prompted to evaluate how well certain traits reflect them, they often exhibit a tendency to endorse those traits if they are socially desirable (Edwards, 1957). As early as 1953, Edwards expressed skepticism regarding the accuracy of item scores on personality assessments, questioning whether respondents' answers genuinely reflected their personal attributes (Edwards, 1953). Furthermore, Edwards' research from that period demonstrated a direct correlation between the likelihood of endorsing an item and its level of social desirability. This inclination toward response bias may stem from various factors such as the experimental or testing environment, the motives of the subjects (e.g., aspirations for achievement, the desire for approval, etc.), or the individuals' anticipation of the evaluative outcomes of their actions (King & Bruner, 2000).

Concerning the impact of Social Desirability (SD) bias, it stands as one of the most prevalent sources of bias affecting the credibility of research findings within psychology and the social sciences (King & Bruner, 2000; Malhotra, 1988; Nederhof, 1985; Paulhus, 1991; Peltier and Walsh, 1990). When relying on self-reported data featuring socially desirable responses, it can lead to false associations between variables, potentially obscuring or weakening the relationships among the variables of interest (Connelly & Chang, 2016; Ganster et al., 1983; Kaiser et al., 2008; Paunonen & LeBel, 2012). Moreover, such bias can skew the average scores on trait questionnaires (Ziegler et al., 2007) and alter the internal structure of measurement instruments (Pettersson et al., 2012). Thus, it is strongly advised to implement methods for controlling or mitigating the influence of SD in research endeavors.

## How to Represent Social Desirability

In the SD literature, there is an ongoing debate on the dimensionality of SD. The single-factor model has been challenged by two-factor models, suggesting that SD consists of two different (but related) factors (e.g., Paulhus, 1984). For example, Paulhus (1984, Paulhus & John, 1998) has presented evidence for the two-factor model of SD, where one factor is labeled impression management and the other is self-deception. More specifically, Paulhus and John (1998) state that SD consists of two self-favoring tendencies: 1) Alpha: an egoistic tendency to see oneself as an exceptionally talented and socially prominent member of society; 2) Gamma: a moralistic tendency, the view of oneself as an exceptionally good member of society. Ziegler et al. (2011) present an argument for the one-factor model. The authors state that it is necessary to show a correlation between scales to introduce a method factor such as SD; if there is no correlation then there is nothing to explain the importance of SD since it won’t affect the correlations between other constructs. They also state that there can be method factors on a more specific level, however, scholars are often concerned only with factors influencing instruments in general, and the single-factor model is often enough for this (Ziegler et al., 2011).

## Modeling Faking with Classical Test Theory

Since faking is a measurement issue, it’s a necessary task to conceptualize faking within the psychometric theory. In a Classical Test Theory perspective, an individual’s observed score ($X$) on a test can be expressed as a function of the person’s true score ($T$) and error ($E$), such that

$$
X = T + E
$$

Then, in a set of observed scores for a sample of test takers, the variance in the observed scores can be expressed as a function of the variance in the true scores and the variance of the errors. Note that, in the equation below, there is an assumption that the error is random and unrelated to true scores. Then, when incorporating faking into the equation, the observed scores associated with faking cannot be due to random error. In other words, faking must be conceptualized as a component of a psychological true score (Ziegler et al., 2011).

$$
\sigma^2_X=\sigma^2_T+\sigma^2_E
$$
In a psychometric approach, it’s common to conceptualize faking as a single, unitary source of systematic variance (e.g., Komar et al., 2008; Schmitt & Oswald, 2006). However, as stated in Ziegler et al. (2011), conceptualizing faking as a single source of systematic variance is an oversimplification, because it is a complex behavior, and the degree to which one fakes is a function of dispositional, attitudinal, and situational factors. In a motivating setting (where people will fake), we can express the observed scores as follows in the following equation:
$$
X_{Motivated}=(T_T+(T_{F1}+...+T_{Fn}))+E
$$
where $T_{F1}$ to $T_{Fn}$ are systematic individual attitudinal, and situational factors that influence observed scores in motivating contexts. In a sample of scores, we can express the variance in observed scores obtained in motivated settings as follows in the equation:
$$
\sigma^2_{X Motivated}=\sigma^2_{T_t}+(\sigma^2_{F1}+...+\sigma^2_{Fn})+(2\sigma^2_{T_T,F1}+...+2\sigma^2_{Fn-1,Fn})+\sigma^2_E
$$

## Some Ways to Control Social Desirability

### Correlations and Social Desirability Scales

Controlling SD bias has been a challenge in the literature (Leite & Cooper, 2009; Paulhus,1981; Ziegler et al., 2011). One of the basic forms of control is to use scales that measure 'pure' desirability, i.e., that measures an SD construct independent of specific content. With this scale, it is possible to measure whether there is a correlation between other instruments and the latent variable of desirability (e.g., Greenblatt et al. 1984). These instruments describe socially desirable but statistically infrequent behaviors. For example, “I don’t gossip about other people’s business” or “I always obey laws, even if I am unlikely to get caught” (Paulhus, 1991). Then, it is interpreted that, if individuals agree with such statements, they lie because it’s virtually impossible to do so. SD scales have been used by many studies across disciplines. Thus, individuals’ scores on these scales are interpreted as indicating how positively a person wants to present themselves.

Much research was conducted using SD scales, for example, in public health, medicine, criminology, and politics (e.g., Hebert et al., 1997; Ng et al., 2020; Vecina et al., 2016; Williams et al., 2009). However, this method assumes that the scale only assesses desirability in isolation (orthogonal) to the target construct (i.e., the construct the researcher wants to measure), that is, these scales fail in terms of the discriminant validity of other constructs related to desirability, since desirability is related to other constructs. In addition to the limitations of using an SD instrument, there is an ongoing debate about what these scales measure (e.g., Connelly & Chang, 2016; Tourangeau & Yan, 2007). For example, some argue that they assess socially desirable personality traits (also called “substance”). Connelly and Chang (2016) provided evidence that these instruments contain both method variance (i.e., response style) and trait variance (i.e., substance). Another meta-analysis showed that SD scales do not measure what they intend to measure (i.e., a positive self-representation; Lanz et al., 2021). This claim comes from the fact that SD scale scores have close to zero correlation with prosocial behavior, even in high-stakes settings (i.e., using monetary incentives, and more anonymous vs. less anonymous research; Lanz et al., 2021).

Another limitation is the validity of SD scales as “faking detectors”. de Vries and colleagues (2014) and Uziel (2010) argue that scores on the SD scale reflect substantive (socially desirable) traits rather than a general response bias. Tourangeau and Yan (2007) state that the key limitation of these scales is the interpretation of the scores. More specifically, it is impossible to differentiate between a truly honest respondent who is virtuous (e.g., people that don’t gossip and always obey the law) and a dishonest respondent who actively lies to present themselves positively.

Meta-analyses have shown that controlling for SD does not increase the predictive validity of scales (Li & Bagger, 2006; Ones et al., 1996). Nonetheless, the studies included in these meta-analyses are based on correlation coefficients and control desirability using partial correlation. These methods have strong assumptions about the psychometric properties of scale items (Leite & Cooper, 2009). To avoid this, newer and more robust methods can be used.

### Ferrando et al. (2009)

Ferrando and colleagues’ (2009) model aims to control biases, including social desirability. The model uses the followingequation (summarized here):

$$
X_{ij}=⍺_{jc}𝛉_{ic} + ⍺_{jd}𝛉_{id} + 𝛆_{ij}
$$

where $⍺$ represents the factor loading, $\theta_{ic}$ represents the content factor (the construct to be measured), $\theta_{id}$ represents the social desirability factor, and $𝛆$ represents the residue/ error. Considering that we are going to use a social desirability scale to measure $\theta_{id}$, we have that, for each $k$ number of items on a SD scale, the model reduces to.

$$
X_{ij}=⍺_{kd}𝛉_{id} + 𝛆_{ij}
$$

Using the framework proposed by Ferrando et al. (2009), model adjustment is made using *minimum rank* factor analysis (MRFA; Ten Berge & Kiers, 1991). Following the requirements of this analysis, MRFA minimizes common variance when multiple r-factors are maintained. To estimate the social desirability factor, it is expected that, in a good test, this factor will be weaker than the other construct to be measured. Therefore, to obtain a stable solution with the present method, the authors suggest having at least three markers of social desirability. The first item is used as a proxy for SD, while the remaining items are taken as instrumental variables.

Ferrando and colleagues' (2009) Semi-Restricted Three-Dimensional Factor-Analytic Model aims to control for acquiescence and social desirability. The model proposed by the authors has two strong assumptions, but only one is of interest for desirability. The assumption states that we have at least one item that measures “pure” social desirability (a proxy variable). On the one hand, some researchers believe that there is no a priori reason why desirability should be related to other latent traits (e.g., Edwards, 1967). On the other hand, social desirability is expected to be related to personality traits such as conscientiousness, emotional stability, agreeableness, and socialization (e.g., Connelly & Chang, 2016; Graziano & Tobin, 2002). Therefore, Ferrando's model does not completely solve the problem of a “pure” social desirability scale, as it requires at least one desirability item completely orthogonal to the psychological dimension.

Another limitation of Ferrando et al. (2009) is that the method only applies to unidimensional scales or with multidimensional measurements that approximate an independent cluster structure. Furthermore, there is a tendency for this method to overcontrol (Uziel, 2010), that is, it tends to overestimate the importance of the bias. This is because in self-report instruments it is common to have a general content dimension (for example, the general kindness factor, which brings together several facets). When estimating a general dimension such as bias, part of that general dimension can be attributed to the true variance of the descriptive content (psychological dimension). Thus, the more orthogonal the desirability factor is in relation to the other latent trait, the lower the chance of excess control occurring. Therefore, control methods within its scale can alleviate this limitation.

### Leite and Cooper (2010)

This model uses factor mixture models as an extension of Ferrando's (2005) method for detecting SD bias. Their method has two contrasting hypotheses: the null hypothesis states that the SD bias factor is not related to subjects' responses on the content scale; the other states that the SD bias factor predicts responses to the focal items for all respondents. Regarding the extent of their work, Leite & Cooper's (2010) method is an intermediate hypothesis: for some individuals in the sample (but not for all individuals), the SD bias factor predicts responses to the items of the content instrument. Although this method can differentiate between SD responders and nonresponders, it still does not solve the problem of a 'pure' SD instrument, because there is still a need to use SD instruments that have 'pure' social desirability items.

### Ziegler and Buehner (2009)

The authors conclude that *faking* can be understood as a systematic measurement error, resulting from the interaction between context and person. If *faking* were seen as this interaction, it would then be systematic variance (Ziegler & Buehner, 2009). Thus, spurious measurement errors are systematic because it is assumed that this error does not always occur, but always occurs in identical circumstances.

Ziegler & Buehner (2009) propose a new way to separate trait variance from *faking*, stating that it is a method to control SD. The logic behind this modeling is that spurious measurement error (i.e., *faking*) contributes to correlations between instruments, however, not between scales that measure a latent trait, but between scales that contain *faking*. Thus, a systematic measurement error can be viewed as common method variance (CMV; Podsakoff et al., 2003), which is modeled as a latent variable using structural equation modeling.

The proposed method works as follows. The questionnaire is administered twice to two groups, and spurious measurement errors must occur at both measurement points in both groups if SD always occurs to some extent. The two groups are separated as 1) a control group, which is asked to answer honestly both times (low stakes); 2) an experimental group, which receives a specific forgery instruction for the second time (high stakes). At the first measurement point, both common factors of the method must have the same character. However, at Time 2, the character of the common method factor in the experimental group should have changed due to the specific *faking* instruction.

One of the limitations of using CMV can be interpreted by Podsakoff et al. (2003), who alerted CMV users, as this latent factor of the method could comprise several things, with *faking* being just one of the explanations. Furthermore, this method assumes that all respondents in a specific condition are *faking* their answers. However, we have no evidence to assume that virtuous people will *fake* their answers in specific scenarios, or that “fakers” will *fake* every time. Another limitation is related to resource constraints, researchers must spend more time, money and other resources collecting more data and ensuring that participants answer the questionnaire twice. The third limitation is stated by the authors:

>To use the presented structural equation model to extract *faking* from variance, at least two different characteristics must be faked by participants. Otherwise, the spurious measurement error variance and the trait variance could not be separated, because the trait and the method factor would attempt to explain the same variance.

### Peabody quadruplets (1967)

To gauge Social Desirability (SD) while preserving the integrity of item content, one approach to managing social desirability is by meticulously crafting items within the scale itself (e.g., Peabody, 1967; Pettersson et al., 2012). This involves dividing items into two distinct components: one addressing the construct under evaluation (descriptive content), and the other focusing on the social desirability of the behavior described by the item (evaluative content). Through this method, balanced quadruplets can be constructed to encompass both the lower and upper extremes of a given characteristic. Consequently, quadruplets offer the additional advantage of mitigating another form of bias known as acquiescence bias – the tendency to endorse items irrespective of their content polarity, whether positive or negative (Mirowsky & Ross, 1991). An illustrative example of a quadruple is presented in the @tbl-peabody.

|                  |  Low Desirability | High Desirability |
|------------------|----------------------|---------------------|
| Low Descriptive  | Withdrawn             | Introspective       |
| High Descriptive | Chatty             | Communicative        |
:Hypothetical Descriptors to Assess Extraversion in Quadruplets {#tbl-peabody}


Peabody's (1967) approach exclusively relies on quadruplets to assess social desirability. However, this method can pose operational challenges. Firstly, the content of the items within the quadruplets may not always lend itself to manipulation or may not naturally fit into a question format. Secondly, employing quadruplets necessitates a substantial number of items to evaluate the same content, which may not contribute additional insights into the construct while potentially inducing respondent fatigue. As a result, one alternative is to estimate social desirability within the quadruplets and utilize this estimation to regulate desirability outside the quadruplets, such as through the employment of multiple indicators multiple causes (MIMIC) modeling techniques.

From the creation of quadruplets it is possible to extract a general factor of social desirability, and then separate the bias from the other factors that we wish to measure (Pettersson et al., 2014; Pettersson et al., 2012; Saucier et al., 2001). The quadruple model can be represented by the following matrix equation:
$$
\begin{bmatrix}
x1\\
x2\\
x3\\
x4
\end{bmatrix}
=
\begin{bmatrix}
-𝛌_{1c} & +𝛌_{1d}\\
-𝛌_{2c} & -𝛌_{2d}\\
+𝛌_{3c} & +𝛌_{3d}\\
+𝛌_{4c} & -𝛌_{4d}
\end{bmatrix}
\begin{bmatrix}
\eta_c\\
\eta_d
\end{bmatrix}
+
\begin{bmatrix}
𝛆_1\\
𝛆_2\\
𝛆_3\\
𝛆_4
\end{bmatrix}
$$

Where $x_n$ is the observed response for item $n$ within the quadruple; $𝛌_{nc}$ is the factor loading of item $n$ in content dimension $c$ ; $𝛌_{nd}$ is the factor loading of item $n$ on the desirability dimension $d$; $\eta$ represents the constructs; and $𝛆$ the measurement errors.

Bastos and Valentini (2023) have run two simulation studies in order to see if controlling the social desirability using the MIMIC model recovers the MIMIC regressions from the social desirability factors to items outside of the quadruplets manipulations. The first simulation showed that, under certain conditions, the MIMIC-Quadruplets model for Likert-type recovered the SD regressions to extra items. In addition, in the MIMIC-Quadruplets model for forced-choice, all conditions simulated in this study recovered (based on bias and coverage indicators) the regressions from social desirability to extra items.

For this, I have developed two intuitive shiny apps where researchers can input their model and see if there’s enough power, low bias, and high coverage to estimate the social desirability of items outside of the quadruplets. One of the apps (called quadSimple; <https://peabody-mimic.shinyapps.io/quadSimple/>) is more user-friendly and requires little information regarding the instrument. The quadSimple is recommended to be used before the construction of an instrument, to give light to the required number of quadruplets they need to build. The other app (called quadSim; <https://peabody-mimic.shinyapps.io/quadSim/>) is more precise and requires more information about the instrument. The quadSim is recommended for scales where researchers already have information on the model parameters.

## How to Control Social Desirability in R

### Controlling Desirability with Ferrando et al. (2009)

To run with the analysis by Ferrando et al. (2009), we first have to install the *vampyr* (Navarro-Gonzalez et al., 2021) package to run the analyses.

``` {.r .R}
install.packages("vampyr")
```

And tell the program that we are going to use the functions of these packages.

```{r}
library(vampyr)
```

To run the analyses, we will use a database from the package itself. Let's see what the dataset looks like.

```{r}
summary(vampyr::vampyr_example)
```

According to the package, we have a dataset with 300 observations and 10 variables, where 6 items measure physical aggression and we have 4 markers of social desirability. Items 1, 2, 3, and 4 are markers of SD ("pure" measures of SD), and the remaining 6 items measure physical aggression. Items 5, 7 and 8 are in the positive pole of the target construct and items 6, 9 and 10 are written in the negative pole of the target construct.

To perform the analysis controlling both desirability and acquiescence, simply use the following code.

```{r}
res <-  ControlResponseBias(vampyr_example,
                      content_factors = 1,
                      SD_items = c(1,2,3,4),
                      corr = "Polychoric",
                      contAC = TRUE,
                      rotat = "promin",
                      PA = FALSE,
                      factor_scores = FALSE,
                      path = TRUE)
```

This analysis allows controlling the effects of two response biases: Social Desirability and Acquiescence, extracting the variance due to these factors before extracting the content variance. If you don't have or want to control acquiescence, simply change the argument `contAC = TRUE` to `contAC = FALSE` .

We see that Bartlett's test of sphericity and KMO were calculated before proceeding with Exploratory Factor Analysis. Furthermore, the model fit indices were calculated. We also see that items 6, 9 and 10 have even high loadings on the desirability factor ("Factor SD"), and items 5, 7 and 8 on the acquiescence factor ("Factor AC").

The cool thing is that it allows you to calculate people's factor scores. Factor scores work like when you calculate the mean score of an instrument to correlate with others, but calculating mean scores has certain assumptions, while factor scores have others. So, to calculate the factor scores while controlling the SD and acquiescence biases, simply leave the factor scores argument as `TRUE` (`factor_scores = TRUE)` and save the result in some variable. In our case, we save the results in the `res` variable.

To save only the factor scores, simply extract the scores from the list.

```{r}
factor_scores <- res$Factor_scores
```

This way, just put this column of factor scores together with your data (using `cbind()`) and then calculate whatever analysis you want.

### Controlling with MIMIC and Quadruples (Bastos & Valentini, 2023)

To run a MIMIC with Quadruples, we first have to install the *lavaan* (Rosseel, 2012) package to run the analyzes and for database simulation and *semplot* (Epskamp, 2022) package for visualization.

``` {.r .R}
install.packages("lavaan")
install.packages("semPlot")
```

And tell the program that we are going to use the functions of these packages.

```{r, include = TRUE}
library(lavaan)
library(semPlot)
```

Let's simulate the data with quadruplets for us to use.
```{r, include = TRUE}
#Quadruple Factor Loadings on Social Desirability
FactorLoadingsSDQ<- rep(0.3, 16)*c(1,-1,1,-1)

#Quadruple Factor Loadings on the Target Construct
RandomFactorLoadingsQ<-rep(0.7, 16)*c(-1,-1,1,1)

# Factor Loads of the extra item in the Target Construct
set.seed(2021)
RandomFactorLoadings <- round(runif((10), min = .3, max = .8), 3)

# Desirability Regressions for Target Construct items
set.seed(2021)
RandomSDregression <- round(runif((10), min = .1, max = .5), 3)

# Item Thresholds
set.seed(2020)
thld1Vet<-round(runif(26, min=-2, max=.5),3)
thld2Vet<-round(thld1Vet +.5,3)
thld3Vet<-round(thld1Vet + 1,3)
thld4Vet<-round(thld1Vet + 1.5,3)

# Simulated Model
simModel <- paste0("fator1 =~",RandomFactorLoadings[1],"*it1 +",
                         RandomFactorLoadings[2],"*it2 +",
                         RandomFactorLoadings[3],"*it3 +",
                         RandomFactorLoadings[4],"*it4 +",
                         RandomFactorLoadings[5],"*it5 +",
                         RandomFactorLoadingsQ[1],"*sd1 +",
                         RandomFactorLoadingsQ[2],"*sd2 +",
                         RandomFactorLoadingsQ[3],"*sd3 +",
                         RandomFactorLoadingsQ[4],"*sd4 +",
                         RandomFactorLoadingsQ[5],"*sd5 +",
                         RandomFactorLoadingsQ[6],"*sd6 +",
                         RandomFactorLoadingsQ[7],"*sd7 +",
                         RandomFactorLoadingsQ[8],"*sd8\n",
                             
                         "fator2 =~", RandomFactorLoadingsQ[6],"*it6 +", 
                         RandomFactorLoadingsQ[7],"*it7 +",
                         RandomFactorLoadingsQ[8],"*it8 +",
                         RandomFactorLoadingsQ[9],"*it9 +",
                         RandomFactorLoadingsQ[10],"*it10 +",
                         RandomFactorLoadingsQ[9],"*sd9 +",
                         RandomFactorLoadingsQ[10],"*sd10 +",
                         RandomFactorLoadingsQ[11],"*sd11 +",
                         RandomFactorLoadingsQ[12],"*sd12 +",
                         RandomFactorLoadingsQ[13],"*sd13 +",
                         RandomFactorLoadingsQ[14],"*sd14 +",
                         RandomFactorLoadingsQ[15],"*sd15 +",
                         RandomFactorLoadingsQ[16],"*sd16\n",
                             
                         "SD =~", FactorLoadingsSDQ[1], "*sd1 +", 
                         FactorLoadingsSDQ[2],"*sd2 +",
                         FactorLoadingsSDQ[3],"*sd3 +",
                         FactorLoadingsSDQ[4],"*sd4 +",
                         FactorLoadingsSDQ[5], "*sd5 +",
                         FactorLoadingsSDQ[6],"*sd6 +",
                         FactorLoadingsSDQ[7],"*sd7 +",
                         FactorLoadingsSDQ[8],"*sd8 +",
                         FactorLoadingsSDQ[9], "*sd9 +",
                         FactorLoadingsSDQ[10],"*sd10 +",
                         FactorLoadingsSDQ[11],"*sd11 +",
                         FactorLoadingsSDQ[12],"*sd12 +",
                         FactorLoadingsSDQ[13], "*sd13 +",
                         FactorLoadingsSDQ[14],"*sd14 +",
                         FactorLoadingsSDQ[15],"*sd15 +",
                         FactorLoadingsSDQ[16],"*sd16\n",
                             
                         "SD ~~ 1*SD\n",
                         "fator1 ~~ 1*fator1\n",
                         "fator2 ~~ 1*fator2\n",
                         "fator1 ~~ 0*SD\n",
                         "fator2 ~~ 0*SD\n",
                         "fator1 ~~ .3*fator2\n",
                             
                         "it1 ~",RandomSDregression[1],"*SD\n",
                         "it2 ~",RandomSDregression[2],"*SD\n",
                         "it3 ~",RandomSDregression[3],"*SD\n",
                         "it4 ~",RandomSDregression[4],"*SD\n",
                         "it5 ~",RandomSDregression[5],"*SD\n",
                         "it6 ~",RandomSDregression[6],"*SD\n",
                         "it7 ~",RandomSDregression[7],"*SD\n",
                         "it8 ~",RandomSDregression[8],"*SD\n",
                         "it9 ~",RandomSDregression[9],"*SD\n",
                         "it10 ~",RandomSDregression[10],"*SD\n",
                             
                         "sd1 |",thld1Vet[1],"*t1 +", thld2Vet[1], "*t2 +", 
                         thld3Vet[1],"*t3 +",thld4Vet[1],"*t4\n",
                         "sd2 |",thld1Vet[2],"*t1 +", thld2Vet[2], "*t2 +", 
                         thld3Vet[2],"*t3 +",thld4Vet[2],"*t4\n",
                        "sd3 |",thld1Vet[3],"*t1 +", thld2Vet[3], "*t2 +", 
                        thld3Vet[3],"*t3 +",thld4Vet[3],"*t4\n",
                        "sd4 |",thld1Vet[4],"*t1 +", thld2Vet[4], "*t2 +",
                        thld3Vet[4],"*t3 +",thld4Vet[4],"*t4\n",
                        "it1 |",thld1Vet[5],"*t1 +", thld2Vet[5], "*t2 +",
                        thld3Vet[5],"*t3 +",thld4Vet[5],"*t4\n",
                        "it2 |",thld1Vet[6],"*t1 +", thld2Vet[6], "*t2 +",
                        thld3Vet[6],"*t3 +",thld4Vet[6],"*t4\n",
                        "it3 |",thld1Vet[7],"*t1 +", thld2Vet[7], "*t2 +",
                        thld3Vet[7],"*t3 +",thld4Vet[7],"*t4\n",
                        "it4 |",thld1Vet[8],"*t1 +", thld2Vet[8], "*t2 +",
                        thld3Vet[8],"*t3 +",thld4Vet[8],"*t4\n",
                        "it5 |",thld1Vet[9],"*t1 +", thld2Vet[9], "*t2 +",
                        thld3Vet[9],"*t3 +",thld4Vet[9],"*t4\n",
                        "it6 |",thld1Vet[10],"*t1 +", thld2Vet[10], "*t2 +",
                        thld3Vet[10],"*t3 +",thld4Vet[10],"*t4\n",
                        "it7 |",thld1Vet[11],"*t1 +", thld2Vet[11], "*t2 +",
                        thld3Vet[11],"*t3 +",thld4Vet[11],"*t4\n",
                        "it8 |",thld1Vet[12],"*t1 +", thld2Vet[12], "*t2 +",
                        thld3Vet[12],"*t3 +",thld4Vet[12],"*t4\n",
                        "it9 |",thld1Vet[13],"*t1 +", thld2Vet[13], "*t2 +",
                        thld3Vet[13],"*t3 +",thld4Vet[13],"*t4\n",
                        "it10 |",thld1Vet[14],"*t1 +", thld2Vet[14], "*t2 +",
                        thld3Vet[14],"*t3 +",thld4Vet[14],"*t4\n",
                        "sd5 |",thld1Vet[15],"*t1 +", thld2Vet[15], "*t2 +",
                        thld3Vet[15],"*t3 +",thld4Vet[15],"*t4\n",
                        "sd6 |",thld1Vet[16],"*t1 +", thld2Vet[16], "*t2 +",
                        thld3Vet[16],"*t3 +",thld4Vet[16],"*t4\n",
                        "sd7 |",thld1Vet[17],"*t1 +", thld2Vet[17], "*t2 +",
                        thld3Vet[17],"*t3 +",thld4Vet[17],"*t4\n",
                        "sd8 |",thld1Vet[18],"*t1 +", thld2Vet[18], "*t2 +",
                        thld3Vet[18],"*t3 +",thld4Vet[18],"*t4\n",
                        "sd9 |",thld1Vet[19],"*t1 +", thld2Vet[19], "*t2 +",
                        thld3Vet[19],"*t3 +",thld4Vet[19],"*t4\n",
                        "sd10 |",thld1Vet[20],"*t1 +", thld2Vet[20], "*t2 +",
                        thld3Vet[20],"*t3 +",thld4Vet[20],"*t4\n",
                        "sd11 |",thld1Vet[21],"*t1 +", thld2Vet[21], "*t2 +",
                        thld3Vet[21],"*t3 +",thld4Vet[21],"*t4\n",
                        "sd12 |",thld1Vet[22],"*t1 +", thld2Vet[22], "*t2 +",
                        thld3Vet[22],"*t3 +",thld4Vet[22],"*t4\n",
                        "sd13 |",thld1Vet[23],"*t1 +", thld2Vet[23], "*t2 +",
                        thld3Vet[23],"*t3 +",thld4Vet[23],"*t4\n",
                        "sd14 |",thld1Vet[24],"*t1 +", thld2Vet[24], "*t2 +",
                        thld3Vet[24],"*t3 +",thld4Vet[24],"*t4\n",
                        "sd15 |",thld1Vet[25],"*t1 +", thld2Vet[25], "*t2 +",
                        thld3Vet[25],"*t3 +",thld4Vet[25],"*t4\n",
                        "sd16 |",thld1Vet[26],"*t1 +", thld2Vet[26], "*t2 +",
                        thld3Vet[26],"*t3 +",thld4Vet[26],"*t4")

#Simulating the Data
simulatedData <- lavaan::simulateData(model = simModel,
                                       model.type = "sem",
                                       sample.nobs = 4000,
                                       seed = 2024,
                                       return.type = "data.frame",
                                       standardized = TRUE
                                       )

```

In the simulated data, we have items from it1 to it10 (which are items that are not made in quadruple format), items sd1 to sd16 (which are items in quadruple format), and are in the 5-point Likert format. See a summary of the items below.

```{r}
summary(simulatedData)
```

You will get to understand the model better now, when we configure it. We have 26 items, 4 of which are quadruples (16 items), and 10 items outside the quadruples to try to control social desirability. Let's configure the model the way you would with your database, that is, we will place all items (quadruples or not) of a given factor to estimate that factor. For example, items it1 to it4 and items sd1 to sd8 are Factor 1 items, so we will estimate Factor 1 with these items. The same logic applies to Factor 2. To estimate desirability, we will only use the quadruples, given that only in the quadruples we manipulated the items to have desirability. We also have to maintain the content factors (Factor 1 and Factor 2) with a correlation equal to 0 with the desirability factor. This is a necessary step to be able to carry out the calculation, otherwise we will have to estimate more parameters than we have information about. Finally, we will perform a desirability regression for the items that were not manipulated in quadruples, to control for the desirability of these extra items.

```{r}
empiricalModel <- "
              factor1 =~ NA*it1 + it2 + it3 + it4 + it5 + sd1 + sd2 + sd3 + 
              sd4 + sd5 + sd6 + sd7 + sd8

              factor2 =~ NA*it6 + it7 + it8 + it9 + it10 + sd9 + sd10 + sd11 + 
              sd12 + sd13 + sd14 + sd15 + sd16
              
              SD =~ NA*sd1 + sd2 + sd3 + sd4 + sd5 + sd6 + sd7 + sd8 + sd9 +
              sd10 + sd11 + sd12 + sd13 + sd14 + sd15 + sd16
              
              SD ~~ 1*SD
              factor1 ~~ 1*factor1
              factor2 ~~ 1*factor2

              factor1 ~~ 0*SD
              factor2 ~~ 0*SD
              factor1 ~~ factor2
              
              it1 ~ SD
              it2 ~ SD
              it3 ~ SD
              it4 ~ SD
              it5 ~ SD
              it6 ~ SD
              it7 ~ SD
              it8 ~ SD
              it9 ~ SD
              it10 ~SD"
```

E agora, rodaremos a análise da seguinte forma. Como temos itens ordinais, falamos para o programa que os itens são ordinais e usamos o estimador "WLSMV".

```{r}
sem.fit <- sem(model = empiricalModel,
               data = simulatedData,
               estimator = "WLSMV",
               ordered = TRUE
               ) 

summary(sem.fit, 
        standardized=TRUE,
        fit.measures = TRUE
        )
```

We see that the fit index was adequate, all factor loadings were significant and all desirability regressions for the extra items were significant.

To extract factor scores to use for other analyses, simply use the following code.

```{r}
data_with_scores <- lavPredict(sem.fit,
                                type = "lv",
                                method = "EBM",
                                label = TRUE,
                                append.data = TRUE,
                                optim.method = "bfgs"
                                )
```

We see that in the variable `data_with_scores`, the factor scores of each subject were calculated and these scores were added to their database.

Let's see an image representation of the model using the code below.

```{r}
semPlot::semPaths(object = sem.fit,
                  layout = "tree2",
                  rotation = 3,
                  whatLabels = "std",
                  edge.label.cex = 0.5,
                  what = "std",
                  edge.color = "black")
```

## References

Bastos, R. V. S., Valentini, F. (2023). *Simulations for two theoretically sound controls for social desirability: MIMIC and Forced-Choice*. (Publication No. 157.932 B33s). Master's thesis, Universidade São Francisco.

Connelly, B. S., & Chang, L. (2016). A meta‐analytic multitrait multirater separation of substance and style in social desirability scales. *Journal of Personality*, *84*(3), 319-334. <https://doi.org/10.1111/jopy.12161>

Edwards, A. L. (1953). The relationship between the judged desirability of a trait and the probability that the trait will be endorsed. *Journal of Applied Psychology*, *37*(2), 90. <https://doi.org/10.1037/h0058073>

Edwards, A. L. (1957). *The social desirability variable in personality assessment and research*. Dryden Press.

Edwards, A. L. (1967). The social desirability variable: A broad statement. In I. A. Berg (Ed.), *Response set in personality assessment* (pp. 32–47). Aldine.

Epskamp S (2022). *semPlot: Path Diagrams and Visual Analysis of Various SEM Packages' Output*. R package. <https://CRAN.R-project.org/package=semPlot>

Ferrando, P. J. (2005). Factor analytic procedures for assessing social desirability in binary items. *Multivariate Behavioral Research*, *40*(3), 331-349. <https://doi.org/10.1207/s15327906mbr4003_3>

Ferrando, P. J., Lorenzo-Seva, U., & Chico, E. (2009). A general factor-analytic procedure for assessing response bias in questionnaire measures. *Structural Equation Modeling: A Multidisciplinary Journal*, *16*(2), 364-381. <https://doi.org/10.1080/10705510902751374>

Graziano, W. G., & Tobin, R. M. (2002). Agreeableness: Dimension of personality or social desirability artifact?. *Journal of Personality*, *70*(5), 695-728. <https://doi.org/10.1111/1467-6494.05021>

Greenblatt, R. L., Mozdzierz, G. J., & Murphy, T. J. (1984). Content and response‐style in the construct validation of self-report inventories: A canonical analysis. *Journal of clinical psychology*, *40*(6), 1414-1420. <https://doi.org/10.1002/1097-4679(198411)40:6\<1414::AID-JCLP2270400624\>3.0.CO;2-K>

Griffith, R., Malm, T., English, A., Yoshita, Y., & Gujar, A. (2006). Applicant faking behavior: Teasing apart the influence of situational variance, cognitive biases, and individual differences. In R. L. Griffith & M. H. Peterson (Eds.), *A closer examination of applicant faking behavior* (pp. 151 – 178). Information Age.

Hebert, J. R., Ma, Y., Clemow, L., Ockene, I. S., Saperia, G., Stanek, E. J., Merriam, P. A., & Ockene, J. K. (1997). Gender differences in social desirability and social approval bias in dietary self-report. *American Journal of Epidemiology*, *146*(12), 1046–1055. <https://doi.org/10.1093/oxfordjournals.aje.a009233>

King, M. F., & Bruner, G. C. (2000). Social desirability bias: A neglected aspect of validity testing. *Psychology & Marketing*, *17*(2), 79-103. [https://doi.org/10.1002/(SICI)1520-6793(200002)17:2\<79::AID-MAR2\>3.0.CO;2-0](https://doi.org/10.1002/(SICI)1520-6793(200002)17:2%3c79::AID-MAR2%3e3.0.CO;2-0)

Lange, F., & Dewitte, S. (2019). Measuring pro-environmental behavior: Review and recommendations. *Journal of Environmental Psychology*, *63*, 92-100. <https://doi.org/10.1016/j.jenvp.2019.04.009>

Lanz, L., Thielmann, I., & Gerpott, F. H. (2022). Are social desirability scales desirable? A meta‐analytic test of the validity of social desirability scales in the context of prosocial behavior. *Journal of Personality*, *90*(2), 203-221. <https://doi.org/10.1111/jopy.12662>

Leite, W. L., & Cooper, L. A. (2009). Detecting social desirability bias using factor mixture models. *Multivariate Behavioral Research*, *45*(2), 271–293. <https://doi.org/10.1080/00273171003680245>

Li, A., & Bagger, J. (2006). Using the BIDR to distinguish the effects of impression management and self‐deception on the criterion validity of personality measures: A meta‐analysis. *International Journal of Selection and Assessment*, *14*(2), 131-141. <https://doi.org/10.1111/j.1468-2389.2006.00339.x>

Malhotra, N. K. (1988). Some observations on the state of the art in marketing research. *Journal of the Academy of Marketing Science*, *16*(1), 4-24. <https://doi.org/10.1177/009207038801600102>

McFarland, L. A., & Ryan, A. M. (2000). Variance in faking across noncognitive measures. *Journal of Applied Psychology*, *85*(5), 812-821. <https://doi.org/10.1037/0021-9010.85.5.812>

Mirowsky, J., & Ross, C. E. (1991). Eliminating Defense and Agreement Bias from Measures of the Sense of Control: A 2 X 2 Index. *Social Psychology Quarterly*, *54*(2), 127. <https://doi.org/10.2307/2786931>

Navarro-Gonzalez D, Vigil-Colet A, Ferrando PJ, Lorenzo-Seva U, Tendeiro JN (2021). *vampyr: Factor Analysis Controlling the Effects of Response Bias*. <https://CRAN.R-project.org/package=vampyr>

Nederhof, A. J. (1985). Methods of coping with social desirability bias: A review. *European Journal of Social Psychology*, *15*, 263–280. <https://doi.org/10.1002/ejsp.2420150303>

Ones, D. S., Viswesvaran, C., & Reiss, A. D. (1996). Role of social desirability in personality testing for personnel selection: The red herring. *Journal of applied psychology*, *81*(6), 660. <https://doi.org/10.1037/0021-9010.81.6.660>

Paulhus, D. L. (1981). Control of social desirability in personality inventories: Principal-factor deletion. *Journal of Research in Personality*, *15*(3), 383–388. <https://doi.org/10.1016/0092-6566(81)90035-0>

Paulhus, D. L. (1984). Two-component models of socially desirable responding. Journal of Personality and Social Psychology, 46(3), 598–609. <https://doi.org/10.1037/0022-3514.46.3.598>

Paulhus, D. L. (1991). Measurement and control of response bias. In J. P. Robinson, P. R. Shaver, & L. S. Wrightsman (Eds.), Measures of social psychological attitudes: Vol. 1. Measures of personality and social psychological attitudes (pp. 17–59). Academic Press. <https://doi.org/10.1016/B978-0-12-590241-0.50006-X>

Paulhus, D. L., & John, O. P. (1998). Egoistic and moralistic biases in self-perception: The interplay of self-deceptive styles with basic traits and motives. Journal of Personality, 66(6), 1025–1060. <https://doi.org/10.1111/1467-6494.00041>

Peabody, D. (1967). Trait inferences: Evaluative and descriptive aspects. *Journal of Personality and Social Psychology*, *7*(4, Pt.2), 1-18. <https://doi.org/10.1037/h0025230>

Peterson, R. A., & Kerin, R. A. (1981). The quality of self-report data: review and synthesis. *Review of marketing*, 5-20.

Pettersson, E., Mendle, J., Turkheimer, E., Horn, E. E., Ford, D. C., Simms, L. J., & Clark, L. A. (2014). Do maladaptive behaviors exist at one or both ends of personality traits? *Psychological Assessment*, *26*(2), 433-446. <https://doi.org/10.1037/a0035587>

Pettersson, E., Turkheimer, E., Horn, E. E., & Menatti, A. R. (2012). The General Factor of Personality and Evaluation. *European Journal of Personality*, *26*(3), 292-302. <https://doi.org/10.1002/per.839>

Podsakoff, P. M., MacKenzie, S. B., Lee, J. Y., & Podsakoff, N. P. (2003). Common method biases in behavioral research: A critical review of the literature and recommended remedies. *Journal of Applied Psychology*, *88*(5), 879-903. <https://doi.org/10.1037/0021-9010.88.5.879>

R Core Team (2024). *R: A Language and Environment for Statistical Computing*. R Foundation for Statistical Computing, Vienna, Austria. <https://www.R-project.org/>

Raymark, P. H., & Tafero, T. L. (2009). Individual differences in the ability to fake on personality measures. *Human Performance*, *22*(1), 86–103. <https://doi.org/10.1080/08959280802541039>

Riggio, R. E., Salinas, C., & Tucker, J. (1988). Personality and deception ability. *Personality and Individual Differences*, *9*(1), 189–191. <https://doi.org/10.1016/0191-8869(88)90050-5>

Rosseel, Y. (2012). lavaan: An R Package for Structural Equation Modeling. *Journal of Statistical Software*, *48*(2), 1-36. <https://doi.org/10.18637/jss.v048.i02>

Saucier, G., Ostendorf, F., & Peabody, D. (2001). The non-evaluative circumplex of personality adjectives. *Journal of Personality*, *69*(4), 537-582. <https://doi.org/10.1111/1467-6494.694155>

Snell, A. F., Sydell, E. J., & Lueke, S. B. (1999). Towards a theory of applicant faking: Integrating studies of deception. *Human Resource Management Review*, *9*(2), 219–242. <https://doi.org/10.1016/S1053-4822(99)00019-4>

Ten Berge, J. M. F., & Kiers, H. A. L. (1991). A numerical approach to the approximate and the exact minimum rank of a covariance matrix. *Psychometrika*, *56*, 309–315. <https://doi.org/10.1007/BF02294464>

Tourangeau, R., & Yan, T. (2007). Sensitive questions in surveys. *Psychological Bulletin*, *133*(5), 859–883. <https://doi.org/10.1037/0033-2909.133.5.859>

Uziel, L. (2010). Rethinking social desirability scales: From impression management to interpersonally oriented self-control. *Perspectives on Psychological Science*, *5*(3), 243–262. <https://doi.org/10.1177/1745691610369465>

Vecina, M. L., Chacón, F., & Pérez-Viejo, J. M. (2016). Moral absolutism, self-deception, and moral self-concept in men who commit intimate partner violence: A comparative study with an opposite sample. *Violence Against Women*, *22*(1), 3–16. <https://doi.org/10.1177/1077801215597791>

de Vries, R. E., Zettler, I., & Hilbig, B. E. (2014). Rethinking trait conceptions of social desirability scales: Impression management as an expression of honesty-humility. *Assessment*, *21*(3), 286–299. <https://doi.org/10.1177/1073191113504619>

Williams, E. A., Pillai, R., Lowe, K. B., Jung, D., & Herst, D. (2009). Crisis, charisma, values, and voting behavior in the 2004 presidential election. *The Leadership Quarterly*, *20*(2), 70–86. <https://doi.org/10.1016/j.leaqua.2009.01.002>

Zickar, M. J., & Robie, C. (1999). Modeling faking good on personality items: An item-level analysis. *Journal of Applied Psychology*, *84*(4), 551. <https://doi.org/10.1037/0021-9010.84.4.551>

Ziegler, M., & Buehner, M. (2009). Modeling socially desirable responding and its effects. *Educational and Psychological Measurement*, *69*(4), 548-565. <https://doi.org/10.1177/0013164408324469>

Ziegler, M., Maccann, C., & Roberts, R. D. (2011). *New perspectives on faking in personality assessment.* Oxford University Press.

```{=html}
<script src='https://storage.ko-fi.com/cdn/scripts/overlay-widget.js'></script>
<script>
  kofiWidgetOverlay.draw('rafaelbastos', {
    'type': 'floating-chat',
    'floating-chat.donateButton.text': 'Support This',
    'floating-chat.donateButton.background-color': '#5bc0de',
    'floating-chat.donateButton.text-color': '#323842'
  });
</script>
```
